Describe and Explore: Using feature-based time series clustering to identify meaningful structures in intensive longitudinal data

### 2023-01-08 manuscript.tex
:sparkles: lin trnd added

\paragraph{Linear Trend} Linear trends are a common research interest for longitudinal data. ``Do patient symptoms improve?'' or ``Does worker productivity decline?'', are familiar types of research questions. Mathematically, such linear trends are commonly captured using (piecewise) linear regression coefficients.

\paragraph{Nonlinearity} Nonlinear trends are important in two regards. Firstly, nonlinearity is a deviation from linear trends and secondly, the shape and structure of the nonlinear trend. Questions like ``Is the development of anxiety a nonlinear process?'' are mathematically captured using nonlinearity parameters such the bicoherence metrics \citep{cuddy2009}. The structure of nonlinear trends is often mathematically captured using polynomial coefficients or more broadly by how wiggly the line is (e.g., estimated degrees of freedom of GAM spline models; similar to the number of spikes \citealp[]{caro-martin2018}).

>>>>

\paragraph{Linear Trend} Linear trends in time series data refer to the presence of a straight-line development of a variable being measured. This type of trend can be observed when there is a consistent increase or decrease in the data over time \citep{nyblom1986}. For psychological time series, researchers have, for example, pointed out linear trends in interpersonal communications \citep{vasileiadou2014}, and emotion dynamics \citep{oravecz2016}. Linear trends are often the simplest way of assessing whether a psychological theory of change is appropriate \citep{gottman1969}. In empirical practice, linear trends are, thus, a common research interest for psychological time series data: ``Do patient symptoms improve consistently?'' or ``Does worker productivity decline continuously?'', are familiar types of research questions. Mathematically, such linear trends are commonly captured using (piecewise) linear regression coefficients, but also using differential equations \citep{chow2019}.

\paragraph{Nonlinearity} Changes in psychology are not always linear \citep{hayes2007}

Nonlinear trends are important in two regards. Firstly, nonlinearity is a deviation from linear trends and secondly, the shape and structure of the nonlinear trend.

``Some examples of nonlinear DEs have already been proposed in studies of ovulatory regulation (Boker, Neale, \& Klump, 2014), circadian rhythms (E. N. Brown \& Luithardt, 1999), cerebral development (Thatcher, 1998), substance use (Boker & Graham, 1998), cognitive aging (Chow \& Nesselroade, 2004), parent-child interactions (Thomas \& Martin, 1976), dyadic relationships (Chow, Ferrer, \& Nesselroade, 2007); and sudden transitions in attitudes (van der Maas, Kolstein, \& van der Pligt, 2003)." \citep{chow2019}

Questions like: ``Is the development of anxiety a nonlinear process?'' are mathematically captured using nonlinearity parameters such as the bicoherence metrics \citep{cuddy2009}.

The structure of nonlinear trends is often mathematically more difficult to capture \citep{castro-alvarez2022a}. One common way is to capture the trends using polynomial coefficients or more broadly by how `wiggly' the line is (e.g., estimated degrees of freedom of GAM spline models; similar to the number of spikes \citealp[]{caro-martin2018}).

depression most likely non-linear  \citep{hosenfeld2015}. Patients in recovery showed sudden changes in the improvement of depression \citep{helmich2020a}
